Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-06-09 PAPER/main.tex
added to Comparison

The heuristics were implemented in C++ and executed on a computer equipped with an Intel Core i3-2120 processor with a frequency of 3.30GHz. The source code for these procedures used for our computational experiments is available at \cite{hawa2018}. We calculated the average optimality of the heuristics for each value of $\delta$ from 1000 instances, where the optimality of each instance is the ratio of the number of strips in the final solution generated by the heuristic to the theoretical minimum $t$, which was calculated using Equation \ref{eqn:lowerbound}. All individual trials were executed in under 160ms.

Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using different values of $\delta$ and $W$ for $n = 500$ and 1000 respectively. Figures in bold indicate the best average optimality for the given combination of $\delta$ and $W$. Initially, we can see that the optimality ratios tend toward 1 as $\delta$ increases, as the proportion of pairs of score widths that can be packed.

\begin{enumerate}
	\item Opt ratios tend to 1 as $\delta \to 1$, as the problem instances become easier to solve, more feasible alignments of pairs of items, more items per strip, reducing the need for more strips, closer to BPP.
	\item $\delta = 0.0$ $n$ strips required, hence all ratios equal.
	\item $\delta = 1.0$,  MFFD and MFFD$^+$ obtain identical opts, since (described above) instance is equivalent to original BPP, $\tau = 0$, so MFFD and MFFD$^+$ perform in the same fashion as original FFD.
	\item Pattern, PS best for wider range of $\delta$ when $W$ high, and smaller range (only 0.1-0.2) when $W = 1250$.
	\item Conversely, MFFD$^+$ best when $W$ small for wider range of $\delta$.
	\item Note that although PS best for 0.1-0.2, difference is negligible, e.g. $n = 500$, $\delta = 0.2$, difference in opt is 0.00829, translates to less than 2 strips difference on average.
	\item MFFD and MFFD$^+$ able to obtain opts $\leq 1\%$ of $t$, PS is not.
	\item PS never best $\delta \geq 0.8$.
	\item $n = 1000$, $W = 5000$, $\delta = 0.7$, PS is best opt, however PS does not produce a single solution containing $t$ strips, whereas there are 152 instances in which MFFD$^+$ produces solutions containing $t$ strips, i.e. there are at least 152 instances out of the 1000 where $k = t$, and PS cannot obtain any. Therefore, MFFD$^+$ solutions must vary significantly (i.e. some solns that are much worse than optimal), as PS is best opt.
	\item Significant difference in opt, $n = 1000$, $W = 1250$, $\delta = 0.6$, difference between MFFD$^+$ and PS opts is 0.10912 (MFFD$^+$ better), which amounts to an average difference of over 50 strips in solns produced.
	\item Noticable drop in opts between $\delta = 0.4$ and $\delta = 0.5$ when $W$ high, as capacity not main concern, more items can now be placed next to one another, so number of strips reduces dramatically, however when $W = 1250$ there is still a capacity issue, regardless of the increased proportion of possible pairings.
	\item anomaly MFFD $n = 1000$, $W = 1250$, $\delta = 0.3$
	\item harder problems, \cite{falkenauer1996}, triplets, fewer items per strip.
	\item cases when opt $\neq 1$: $t$ incorrect, solution produced is not optimal, or both.
	\item less than 10\% past $\delta = 0.8$
\end{enumerate}

\begin{itemize}
	\item optimality converge toward 1 as $\delta$ increases, i.e. as the problem instances become less difficult.
	\item When $\delta = 1.0$. MFFD and MFFD$^+$ perform identically, as the problems are equivalent to the original BPP, and so they behave like the original FFD heuristic.
	\item When $\delta = 0.0$, all optimalities are equal, as each item requires its own strip, and so $|\mathcal{I}|$ strips are needed for all 1000 instances.
	\item Values in bold represent the best avg optimality of the three heuristics for the given level of $\delta$. 
	\item Clear pattern across both set sizes, PS performs the best out of the three for $\delta = 0.1$ to 0.7 when $W = 5000$, and for $\delta = 0.1$ to 0.6 when $W = 2500$.
	\item However, when $W = 1250$, PS is only superior when $\delta = 0.1$ and 0.2. 
	\item On further inspection, we can see that the difference between the average optimalities for PS and MFFD$^+$ is negligible.
	\item Using Table 1 as an example, when $\delta = 0.2$, the solutions generated by PS are on average within 31.781\% of $t$, and for MFFD$^+$ within 32.61\%, a difference of 0.829\%. 
	\item In practical terms, PS produces solutions using 303.95 strips on average, where $t = 230.648$. MFFD$^+$ produces solutions using 305.862 strips, a difference of less than two strips. 
	\item Another trend is that PS produces average solutions within 10\% of $t$ for $\delta = 0.5$ onwards when $W = 5000$ and 2500, however when $W = 1250$, PS only produces solutions within 10\% of $t$ for $\delta = 0.8$ onwards.
	\item In contrast, MFFD and MFFD$^+$ can produce solution within 10\% of $t$ when $W = 1250$ for $\delta = 0.5$ onwards for $n = 500$, and $\delta = 0.4$ for $n = 1000$.
	\item MFFD$^+$ can also find solutions within 1\% of $t$ for $\delta = 0.8$ onwards for all strip widths when $n = 1000$, and the two larger strip widths when $n = 500$.
	\item Number of items per strip smaller, problem is harder, falkenauer 1996, martello and toth 1990b items per strip roughly 2, triplets hardest problems
	\item all instances run in less than 160ms
	\item For each instance where opt $>1$, two cases: either $t$ is incorrect, and there does not exist a solution with $k = t$ strips, and the solution found is actually an optimal solution, or $t$ is correct, and the heuristic cannot find a solution with $t$ strips, not an optimal solution.
	\item since this is a relatively new problem, no problem instances solved to optimality, and so we have no way of knowing if in fact the heuristics are finding optimal solutions. Only able to compare with the theoretical minimum. 
	\item $W =5000$ Table 2, $\delta = 0.7$ PS has best optimality, however looking at the number of instance solved to $t$, PS has 0, whereas MFFD$^+$ has 152. Therefore there are at least 152 instances where the optimal number of strips = $t$ that PS is not able to solve to optimality.
	\item We see that PS produces solutions using fewer strips on average of the three algoritihms for a larger range of $\delta$ when the strips are wider, and MFFD$^+$ produces solutions using fewer strips for a wider range of $\delta$ when the strips are shorter.
	\item We can see that the range of values of $\delta$ in which PS produces solutions using the minimum number of strips on average of the three algorithms reduces as the strip width decreases
	\item PS never performs best for 0.8 onwards
	\item significant drop between $\delta = 0.4$ and 0.5 for all heuristics and both set sizes except when $W = 1250$
\end{itemize}

@@

>>>>

The heuristics were implemented in C++ and executed on a computer equipped with an Intel Core i3-2120 processor with a frequency of 3.30GHz. The source code for these procedures used for our computational experiments is available at \cite{hawa2018}. We calculated the average optimality of each heuristic for every combination of $n$, $\delta$, and $W$ from 1000 instances, where the optimality $r$ of each instance is
\begin{equation}
	r = \frac{|\mathcal{S}|}{t},
\end{equation}
i.e. the ratio of the number of strips in the final solution $\mathcal{S}$ generated by the heuristic to the theoretical minimum $t$, which was calculated using Equation \ref{eqn:lowerbound}. All individual trials were executed in under 160ms.

Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using different values of $\delta$ and $W$ for $n = 500$ and 1000 respectively. Figures in bold indicate the best average optimality for the given combination of parameters. Initially, we can see that $r$ tends toward 1 as $\delta$ increases, since the proportion of score widths meet the vicinal sum constraint increases and more items can be packed onto one strip, reducing the number of strips required. Note that when $\delta = 1.0$ MFFD and MFFD$^+$ have identical optimality ratios, as the instances are equivalent to the original BPP ($\tau = 0$) and so they operate in the same fashion as the original FFD heuristic.

Looking at results for both $n = 500$ and $n = 1000$, there is a clear pattern with respect to $r$ across all widths and proportion levels. PS has the best optimality ratio for a wider range of $\delta$ when the strips are wider, and a much smaller range when the strip width decreases. Conversely, MFFD$^+$ obtains better optimality ratios for a wider range of $\delta$ when $W = 1250$. Although PS does have the best optimality of the three heuristics for $\delta = 0.1$ and 0.2 using the smallest strip width, we can see that it only marginally superior to MFFD$^+$. For example, take $\delta = 0.2$ in Table \ref{table:n=500} for $W = 1250$. The difference between $r$ for PS and MFFD$^+$ is 0.00829, which translates to less than 2 strips difference between the average number of strips generated by each heuristic.

One interesting feature is the reduction of $r$ between $\delta = 0.4$ and $\delta = 0.5$, which is much more significant for $W = 5000$ as opposed to $W = 1250$. The reduction occurs due to an increase in the number of feasible alignments of the items, as every score width is able to be paired with $n-1$ other score widths when $\delta = 0.5$. Thus, when $W = 5000$, more items can be packed onto a strip feasibly, resulting in fewer strips in the final solution. However, the difference in $r$ is less noticeable for $W = 1250$, as regardless of the increased proportion in possible pairings, the capacity of the strips is much smaller, preventing the collection of items from multiple strips onto one.

\begin{enumerate}
	\item MFFD and MFFD$^+$ able to obtain opts $\leq 1\%$ of $t$, PS is not.
	\item PS never best $\delta \geq 0.8$.
	\item $n = 1000$, $W = 5000$, $\delta = 0.7$, PS is best opt, however PS does not produce a single solution containing $t$ strips, whereas there are 152 instances in which MFFD$^+$ produces solutions containing $t$ strips, i.e. there are at least 152 instances out of the 1000 where $k = t$, and PS cannot obtain any. Therefore, MFFD$^+$ solutions must vary significantly (i.e. some solns that are much worse than optimal), as PS is best opt.
	\item Significant difference in opt, $n = 1000$, $W = 1250$, $\delta = 0.6$, difference between MFFD$^+$ and PS opts is 0.10912 (MFFD$^+$ better), which amounts to an average difference of over 50 strips in solns produced.
	\item Noticable drop in opts between $\delta = 0.4$ and $\delta = 0.5$ when $W$ high, as capacity not main concern, more items can now be placed next to one another, so number of strips reduces dramatically, however when $W = 1250$ there is still a capacity issue, regardless of the increased proportion of possible pairings.
	\item anomaly MFFD $n = 1000$, $W = 1250$, $\delta = 0.3$
	\item harder problems, \cite{falkenauer1996}, triplets, fewer items per strip.
	\item cases when opt $\neq 1$: $t$ incorrect, solution produced is not optimal, or both.
	\item less than 10\% past $\delta = 0.8$
\end{enumerate}

@@

\begin{comment}
	\begin{enumerate}
	\item Opt ratios tend to 1 as $\delta \to 1$, as the problem instances become easier to solve, more feasible alignments of pairs of items, more items per strip, reducing the need for more strips, closer to BPP.
	\item $\delta = 0.0$ $n$ strips required, hence all ratios equal.
	\item $\delta = 1.0$,  MFFD and MFFD$^+$ obtain identical opts, since (described above) instance is equivalent to original BPP, $\tau = 0$, so MFFD and MFFD$^+$ perform in the same fashion as original FFD.
	\item Pattern, PS best for wider range of $\delta$ when $W$ high, and smaller range (only 0.1-0.2) when $W = 1250$.
	\item Conversely, MFFD$^+$ best when $W$ small for wider range of $\delta$.
	\item Note that although PS best for 0.1-0.2, difference is negligible, e.g. $n = 500$, $\delta = 0.2$, difference in opt is 0.00829, translates to less than 2 strips difference on average.
	\item MFFD and MFFD$^+$ able to obtain opts $\leq 1\%$ of $t$, PS is not.
	\item PS never best $\delta \geq 0.8$.
	\item $n = 1000$, $W = 5000$, $\delta = 0.7$, PS is best opt, however PS does not produce a single solution containing $t$ strips, whereas there are 152 instances in which MFFD$^+$ produces solutions containing $t$ strips, i.e. there are at least 152 instances out of the 1000 where $k = t$, and PS cannot obtain any. Therefore, MFFD$^+$ solutions must vary significantly (i.e. some solns that are much worse than optimal), as PS is best opt.
	\item Significant difference in opt, $n = 1000$, $W = 1250$, $\delta = 0.6$, difference between MFFD$^+$ and PS opts is 0.10912 (MFFD$^+$ better), which amounts to an average difference of over 50 strips in solns produced.
	\item Noticable drop in opts between $\delta = 0.4$ and $\delta = 0.5$ when $W$ high, as capacity not main concern, more items can now be placed next to one another, so number of strips reduces dramatically, however when $W = 1250$ there is still a capacity issue, regardless of the increased proportion of possible pairings.
	\item anomaly MFFD $n = 1000$, $W = 1250$, $\delta = 0.3$
	\item harder problems, \cite{falkenauer1996}, triplets, fewer items per strip.
	\item cases when opt $\neq 1$: $t$ incorrect, solution produced is not optimal, or both.
	\item less than 10\% past $\delta = 0.8$
	\end{enumerate}
\end{comment}
